---
title: "Fabs for Nonconvex Loss and Adaptive LASSO"
author: "Xingjie Shi"
date: 2019-03-07T17:13:14-05:00
#categories: ["R"]
tags: ["NSFC 71501089", "high dimension",  "adaptive LASSO"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

## Introduction 
It is common to have multiple outcomes of interest in a single study. Usually, these outcomes are analysed separately, ignoring the correlation between them. In high dimensional setting, one would expect that a carefully designed multivariate approach would be a more efficient alternative to individual analyses of each outcome [cite]. For jointly modeling multiple correlated outcomes, we have developed the method [__vimco__](https://academic.oup.com/bioinformatics/article-abstract/35/19/3693/5372680?redirectedFrom=fulltext). Here we provide same demonstrations.
 
Assume we have measured $q$ responses $y_1, \dots, y_q$ and the same set of $p$ predictors on each individual $x_1,\dots, x_p$. Each response has its own regression model:
$$h(y_i) =  x_i^\top\beta + \varepsilon_i, ~ i=1, \dots, n,$$
where $\beta$ is the regression vector.  


## Simulate high dimensional data
You can simulate a simple example with following R code:

```{r, fig.show='hold'}
AR = function(rho, p) {
    outer(1:p, 1:p, FUN = function(x, y) rho^(abs(x - y))) 
}
library(mvtnorm)
n = 300  # sample size
p = 400  # dimension of predictors
K = 5    # dimension of outcomes
set.seed(20132014)
X   = rmvnorm(n, mean=rep(0, p))
sigma.beta = rep(1, K)

bet = matrix(0, nrow = p, ncol = K)
lambda = 0.15    # proportion of non-zero entries in matrix \beta
eta = rbinom(p, 1, lambda)
alpha = 1
gam = matrix(rbinom(p*K, 1, alpha), ncol=K)
for (k in 1:K){
  bet[, k] = rnorm(p, mean = 0, sd = sigma.beta[k]) * gam[,k] * eta 
}
  
sigma = AR(0.8, K)
lp  = X %*% bet
sigma.e = diag(sqrt(diag(var(lp)))) %*% sigma %*% diag(sqrt(diag(var(lp))))
err = rmvnorm(n, rep(0, K), sigma.e)
Y   = lp + err
print(round(cor(Y),2))
```



## Fit each outcome seperately
```{r, fig.show='hold'}
library(vimco)
fit_Ind = emInd(X, Y)
```
 